Inefficient Concessions and Mediation

### 2021-01-25 draft.tex
Copy edit Sections 2 through first half of 5

Countries\footnote{Players will be referred to as countries, though it could also be appropriate to think of players as intra-state actors, as in a civil war. } are of two possible types, high and low. A high type discounts future period payoffs at rate $\delta_h$ and a low type discounts at $\delta_l$ where $\delta_h >\delta_l$ and $\delta_h ,\delta_l$ $\in [0,1].$ Before the start of the game, nature independently determines the types of country 1 and country 2. The type distinction, as explained in detail later, encompasses the country's willingness to cooperate in solving the conflict; that is, throughout the analysis we will define a low type as a player whose only sequentially rational strategy is to remain in conflict in all periods. It is not necessary to think of a type as fixed; it can change as regimes and relations between countries change.\footnote{We can enrich the model to assign a probability (and even different probabilities) of a country's type shifting from high to low or low to high in a given period.}

 The probability of nature selecting the high type for a given country is $p$. The probability of nature selecting the low type is $1-p$.\footnote{Note that parameters are symmetric across countries, as is consistent with the literature. Qualitative results do not depend on this being the case, and equilibria and equilibrium concessions can easily be calculated with asymmetric parameters between players.}  Countries are aware of their own type, but not the type of the other country.  Similar scenarios have been modeled as credible commitment problems in the bargaining literature. Our model can be thought of as a reduced form bargaining model, but it can also be considered somewhat differently -- when trust is the issue to be solved, there are several solutions to conflict. Concessions are one way in which countries can signal their commitment.  By modeling the conflict in such a way we can also gain insight as to how a third party can ameliorate trust issues inherent to the conflict. %might want to cite fearon and bargaining lit here
%might want to eventually add the stochastic type switch- or at least talk about how that changes analysis

@@

Here $T>0$ represents the benefit from the other country playing Trust. $D>0$ represents the damages incurred when the other country plays Fight, while $W>0$ is the additional benefit a country receives when it plays Fight itself. We assume $T>W-D$ in line with the assumption that the countries play a Prisoner's Dilemma in the stage game.

We will refer to the stage-game Nash Equilibrium of (Fight, Fight) as ``War.'' We will refer to (Trust,Trust) as ``Peace.'' The assumption $T>W-D$ ensures that the joint payoffs from ``Peace'' are the highest among all the stage-game action profiles.

@@

We will use grim trigger punishments throughout: that is, (Fight, Fight) will be played forever if Fight is played by any party in any round. Because we use grim trigger punishments, there are only 4 interesting paths of play after concessions are given:
1) Both countries play Trust always, 
2) Both play Fight always, 
3) Country 1 plays Fight in round 1 while country 2 plays Trust, which is followed by both countries playing Fight from round 2 onwards and 
4) Country 2 plays Fight in round 1 while country 1 plays Trust, which is followed by both countries playing Fight from round 2 onwards.

Equilibrium payoffs beginning from round 1 are now easily described by noting only the collective first round behavior. Let $X_{ij}$ represent the sum of discounted payoffs for country $i$ where subscripts represent first round strategies of that country. If both countries play T in round 1, the corresponding payoff to country $i$ is represented as $ X_{TT}=\frac {T}{1- \delta_i}$. Likewise $X_{TF}=({-D}+\frac{\delta_i(W-D)}{1- \delta_i} )$, $ X_{FF}=\frac{W-D}{1- \delta_i}$, and  $X_{FT}=T+W +\frac{\delta_i(W-D)}{1- \delta_i}$. Note that because types have different discount rates, the payoffs differ by type, and when ambiguous will be super-scripted to indicate the type (e.g. $X_{FT}^h$ represents payoffs for the high type of country 1 when country 1 fights and country 2 trusts in the first period).

@@

\section{Analysis of Benchmark Cases}
\label{sec:bench}
When countries are faced with uncertainty about each other's motives, they can proceed in several different ways, including giving concessions in order to signal their willingness and ability to make peace. We will examine three types of equilibria.\footnote{If the discount factor is high enough, there are many types of equilibria. Consider for instance oscillating every other period between (Trust,Trust) and (Fight,Fight) with a grim trigger War threat. In this paper, we are not interested in these equilibria because they are not as realistic for applications to conflict scenarios. These are also always payoff dominated by at least one of the other equilibria discussed and thus 	are not attractive.} The Appendix provides a thorough analysis of the various equilibria as well as most proofs for this section.

\subsection{Pooling equilibrium}
In the \emph{pooling equilibrium}, both countries do not give concessions in period 0, then always play Fight in every period beginning with period 1. 
Countries continue to doubt, doing nothing to bridge the credibility gap, and thus not cooperating. This \emph{pooling equilibrium} is likely to happen when countries are in long-set patterns of distrust. Active fighting is not necessary; remaining outside of a state of peace is all that is required.  See the Appendix for a formal characterization of this type of equilibrium.

@@

Without a third party, there is little hope of overcoming mistrust. These long lasting stalemates are the types of conflict which are most appropriate for mediator involvement. A mediator will be added to the model later to illustrate exactly how mediation can help achieve peace in this context. In some sense, these are the cases that are most interesting because the international community might play a positive role.

\subsection{No-Concessions Separating Equilibrium}
\label{sec:ncse}
The other two equilibria on which we focus are separating equilibria in which low types and high types have different strategies. If $\delta_i$ is high enough for both countries, the ``Peace'' outcome can be sustained with no concessions and a grim trigger punishment threat. We call this a \emph{no-concessions separating equilibrium}. In this equilibrium, no concessions are given by either type in period 0. In period 1, high types play Trust while low types play Fight. So long as there are two high types, as revealed by their period 1 play, both countries continue to play Trust in all following periods. Otherwise, both countries play Fight in all following periods.

@@

\begin{theorem}
\emph{No Concessions Separating Equilibrium}\\
	Assume $\delta_h \geq \frac{W}{(1-p)W + p(T+D)}=\de^{nc}$. A subgame perfect Nash equilibrium exists in which low types give no concessions in period 0 and always play Fight and the high types give no concessions in period 0 and play Trust unless the other country defects to Fight.
	\label{theorem:1}
\end{theorem}
\emph{Proof}: See Appendix.
\\
Here, as in each further iteration of the model, we define a high type as a country who prefers to play the ``Peace'' equilibrium, i.e. $\delta_h \geq \delta^{nc}$. Any country with $\delta_l < \delta^{nc}$ is a low type. This means that two high types---and only two high types---can sustain the Peace outcome in a simple repeated game.

In this equilibrium, two mutually distrustful countries behave as if they trust anyway, taking a risk. This generally will happen when the result of trusting and then being betrayed is not catastrophic, or the odds of such a scenario are small. This is representative of a case where countries have suspicions about each other, but proceed anyway.

Examples of a \emph{no-concessions separating equilibrium} are most likely to be those where countries have traditionally had good relations but something has recently come in the way. Certain relationships between the USA and countries that its wikileaks alienated -- such as Italy -- would be an example. Italy essentially laughed off the criticisms and reassured the USA of its commitment to their relationship. Warsaw Pact countries generally allowed the USSR similar leeway; it was better to have occasionally imperfect cooperation than conflict. Some events can disrupt trust without completely derailing cooperation.

@@

%Though it will not be modeled here, a stochastic component -- when countries occasionally unilaterally defect -- can explain why powerful countries like the USSR more often ``get away'' with cheating: the payoffs to the relationship are not symmetric. Losing the USA or USSR as an ally is much more devastating for a satellite country than vise versa, therefore big countries can have their alliances and take advantage of them too.
\subsection{Concessions Separating Equilibrium}
\label{sec:cse}
The other separating equilibrium of interest involves distrustful countries to using concessions as a way of building trust. We call this the \emph{concessions separating equilibrium}. In this equilibrium, the cost of giving a concession can be worth the investment if the other country also cooperates. It is also not as risky as the \emph{no-concession separating equilibrium}. The concessions are given before more substantive actions, thus concessions are a costly signal that the other country is serious about cooperating. Most agreements come with concessions, and in many cases, though not all, they are used as signals. The goal is to use this option to avoid continual distrust. That is, high types make use of concessions to identify themselves as trustworthy partners in hopes of establishing Peace.

@@

Details of the analysis, which uses standard incentive compatibility constraints, are in the Appendix. We begin by characterizing the concession (or \emph{gift}) giving behavior of the low type. We then present the patience threshold and gift giving behavior of the high type.

@@

A large probability of a high type makes the \emph{no-concessions separating equilibrium} very attractive because the risk of ending up in the cheater's corner of the Prisoner's Dilemma is low and no concessions have to be made. Conversely, for smaller values of $p$, high types are better off paying the gift to avoid the possibility of being taken advantage of by a low type.

@@

\subsection{Concessions with Material Value}
We now model concessions to have real future material value that can either help or harm the giver. To be clear, in Sections~\ref{sec:model} and \ref{sec:bench}, the gift only benefits the party who receives it in the period in which it is received, that is, Period 0. In this section, we assume that this is an additional value from the gift in each period that the repeated game is played, that is, Period 1 and onward.

The interpretation is that the country which receives the concessions can use them to build up their military or to build up their civil society. Using concessions to build military strength hurts the giver of the concessions in war; using the concessions to strengthen civil society helps the giver in peace. The Golan heights is such an example: if Syria is an ally of Israel, the Golan Heights is a reasonable concession to make, but if Syria later turns out to betray a trust, surrender of the strategic land would be disastrous for Israel. %Another example is ADD more examples

@@

Because $\al_i$ enters into the payoffs in a linear fashion, the choice that maximizes welfare is $\alpha=1$ for high types since a concession would only come from another high type in equilibrium, and thus only the ``Peace'' outcome will be played. Conversely, in equilibrium, low types only participate in the ``War'' equilibrium and thus maximize their payoffs by choosing $\alpha=0$.

@@

$\de_h$ is not involved in any of these calculations, so we can see that for any $\de_h \in [.5\overline{3},1)$\footnote{Recall that we restrict $p<1$ since there is no imperfect information when only one type is possible.} Thus, separating through concessions is possible when concessions have no material value but is not possible when they do have future material value. \hfill $\blacksquare$

@@

We implement this by allowing countries to choose a scalar $e$ to multiply by their given concession $g$ where $0 \leq e \leq 1$. As before, $C(g)= g$ but while the receiving country observes the full concession $g$, the future benefit to the recipient is now only $eg$.\footnote{In a more general model, we could also think of $e$ as a function $E(\cdot)$. The argument is still $g$, but instead of a scalar, $E(\cdot)$ can transform $g$ into any concession. In particular, this allows the transformation to increase the value of the concession. The transformation function $E(\cdot)$ can be thought of as the type of gift, and has a cost to the sender and a direct benefit to the receiver in addition to its signaling value.} Table 3 provides the modified payoffs in the repeated game.

@@

The intuition for Lemma~\ref{lemma:efficiency} is that future material value of concession helps you when your negotiating partner turns out to be a high type, and this happens $p$ of the time. On other other hand, the material value of your concession \emph{hurts} you when your negotiating partner turns out to be a low type. Similarly, the larger are the benefits from Trust, the larger the welfare gain when the negotiating partner is a high type, and the larger are the damages from Fight, the larger is the welfare loss when the negotiating partner turns out to be a low type.

Lemma~\ref{lemma:efficiency} implies that the high types will either prefer $e=1$ or $e=0$. As we turn to our result about the incentive compatibility of the \emph{concessions separating equilibria with material value}, we focus on these two cases going forward as the most focal and easiest upon which to coordinate.\footnote{For some parameter values, we cannot rule out that a \emph{concessions separating equilibrium} with $0<e<1$ exists and improves upon high type welfare when high types are patient enough to separate when $e=0$ but not when $e=1$. This is not at all obvious though, especially because increasing $e$ from $0$ can tighten the high type's incentive constraint.}

@@

This means that there are parameters under which Peace cannot be achieved in a \emph{concessions separating equilibrium with efficient gifts} but Peace can be achieved in a \emph{concessions separating equilibrium with inefficient gifts}. This is one of the reasons that Theorem~\ref{theorem:4} is true.

\begin{theorem}
		Assume the high types are not patient enough to separate without concessions and that concessions have material value that can be destroyed. There are parameters under which the optimal equilibrium is a separating equilibrium in which concessions are inefficient.
	\label{theorem:4}
\end{theorem}
\emph{Proof}: We invoke Lemmas~\ref{lemma:efficiency} and \ref{lemma:e-patient}. There are two reasons for which the parties would choose the \emph{concessions separating equilibrium with inefficient gifts}. First, they may be patient enough to separate when gifts are inefficient but not when gifts are efficient (Lemma~\ref{lemma:e-patient}). In this case, the \emph{concessions separating equilibrium with efficient gifts} is not available to them.

@@

Inefficient concessions are optimal under low $p$ (there's a good chance of facing a low type who will use the concession against you) as well as when the benefits from Peace are low relative to the damages from War. This finding of inefficiency is for profoundly different reasons than the literature related to Camerer (1988) or Prendergast and Stole (2001).\footnote{Since our initial setup is essentially identical to Camerer's, inefficiency of  pre-play communication can also be found in our model. With a slight behavioral modification, the reasoning of Prendergast and Stole (2001) can also be had: if a country suffers regret after efficiently giving to a type different than their own, there is a premium for guessing correctly. In Prendergast and Stole the two types are equal except in labeling; in our model, high types and low types are used so there is an incentive for low types to try to pose as high types. However, putting this difference aside, the inefficiency result can be viewed as similar to that in Prendergast and Stole (2001). That is, if countries suffer a ``mismatching'' cost due to giving to the wrong type, they will be hesitant to do so and may offer inefficient concessions.} Instead of being in reaction to a behavioral regret for mismatching or serving as a way to discourage low types for entering the game altogether, here making concessions inefficient is a device to prevent low types from using those concessions against oneself in the future.

@@

In Sections~\ref{sec:bench} and \ref{sec:IC}, the same concession must be sent to both types since types are private information. Importantly, here the mechanism $M$ can differentiate based on recipient's type. Otherwise, the setup is the same as in Section~\ref{sec:matval} where concessions have future material value that can optionally be destroyed. As usual, truthful self-revelation requires that the individual rationality and incentive compatibility constraint for each type must be satisfied.

@@

\begin{lemma}
	A manipulative mediator cannot facilitate the most efficient concessions---that is, $e=1$. 
	\label{lemma:med}
\end{lemma}
\emph{Proof}: See Appendix.

@@

\begin{theorem}
	Assume that concessions have future material value.
	\begin{enumerate}
		\item[(a)] If future material value can be destroyed, a mediator can be used to eliminate inefficient concessions and improve welfare.
		\item[(b)] If future material value cannot be destroyed, a mediator can be used to expand the parameter space over which peace can be achieved.
	\end{enumerate}
	\label{theorem:5}
\end{theorem}
\emph{Proof:} (a) Under $M$, the low type incentive constraint can bind without having high types give a concession to low types. There is no need then to give inefficient concessions: high types will invest in civil society and improve the giver's welfare and low types will never receive the concession. %The equilibrium separating concession $g^{*M}$ will be larger than $g^*$ in this case because the low type who reports truthfully receives a concession with zero probability: 
%\be
%g^{*M}= \frac{p(X_{FT}^l-X_{FF}^l)}{1-p}
%\ee
High type utility under $M$ is 
\be
U_h^M= pX_{TT}+(1-p)X_{FF} = p \left( \frac{T(1+g^{*M})}{1-\de_h} \right) + (1-p) \left(\frac{W-D}{1-\de_h} \right)
\ee
because high types exchange concessions of the same value. Welfare, defined as high-type utility, is improved because no transfer is made to the low types and no future damage is caused in the case of such a transfer.\footnote{Low types IR constraint still binds and $U_l=X_{FF}$.}

@@

The low type leader can bluff in the bilateral setting but not when a mediator is present. As per Theorem~\ref{theorem:4}, inefficiency is due to the high types giving an inefficient gift in order to separate while still being able to prevent low types from using concessions against the high types. But because the benefits of giving a concession differ between the types (low types get one period of cheater's payoff, while high types get future peace), the mediator is able to mandate concessions only between declared high types. This means there is no longer a need for inefficient concessions since concessions will never be used against the giver. Thus, a mediator is able to remove inefficiency in concession giving and bring willing parties to the table to give the gifts necessary to achieve peace.

@@

 \noindent{\emph{Proof of Theorem \ref{theorem:1}}:\\
Playing Trust in the first period and thereafter as long as Trust has always been played will be incentive compatible for high types if the expected payoffs from this strategy are greater than the payoffs from playing Fight in the first period and then every period thereafter by the one-shot deviation principle. If both countries play Trust in a period, a Peace equilibrium takes place and both countries continue to play peace in every period with the grim trigger threat of the War equilibrium. Low types play Fight in the first period, so a high type that observed Fight being played would respond by playing Fight in period 2 and in all future periods. A country plays Trust if and only if the following condition is satisfied:
\begin{equation*}
  p\left(\frac{T}{1- \delta_i}\right) +(1-p)\left({-D}+\frac{\delta_h(W-D)}{1- \delta_i} \right)\geq p \left(T + W + \frac{\de_h}{1-\de_h}(W-D)\right) +(1-p)\frac{W-D}{1- \delta_h}
\end{equation*}
Simplifying, we have
%\be
%  p\left(\frac{T}{1- \delta_h}\right) \geq p \left(T + W + \frac{\de_h}{1-\de_h}(W-D)\right) +(1-p)W
%\ee
\begin{equation*}
  pT \geq p \left((1-\de_h)(T + W) + \de_h (W-D)\right) + (1-\de_h)(1-p)W
\end{equation*}
\begin{equation*}
  pT \geq pT -p\de_hT +pW -p\de_h W + p \de_h W -p \de_h D  + W -\de_h W -pW + p\de_hW
\end{equation*}
\begin{equation*}
  p\de_hT +p \de_h D + \de_h W - p\de_hW  \geq W 
\end{equation*}
\be
	\delta_h \geq \frac{W}{(1-p)W + p(T+D)}
	\label{eq:ncse}
\ee

>>>>

Countries\footnote{Players will be referred to as countries, though it could also be appropriate to think of players as intra-state actors, as in a civil war.} are of two possible types, high and low. A high type discounts future period payoffs at rate $\delta_h$ and a low type discounts at $\delta_l$ where $\delta_h >\delta_l$ and $\delta_h ,\delta_l$ $\in [0,1].$ Before the start of the game, nature independently determines the types of country 1 and country 2. The type distinction, as explained in detail later, encompasses the country's willingness to cooperate in solving the conflict; that is, throughout the analysis we will define a low type as a player whose only sequentially rational strategy is to remain in conflict in all periods. It is not necessary to think of a type as fixed; it can change as regimes and relations between countries change.\footnote{We can enrich the model to assign a probability (and even different probabilities) of a country's type shifting from high to low or low to high in a given period.}

The probability of nature selecting the high type for a given country is $p$. The probability of nature selecting the low type is $1-p$.\footnote{Note that parameters are symmetric across countries, as is consistent with the literature. Qualitative results do not depend on this being the case, and equilibria and equilibrium concessions can easily be calculated with asymmetric parameters between players.}  Countries are aware of their own type, but not the type of the other country.  Similar scenarios have been modeled as credible commitment problems in the bargaining literature. Our model can be thought of as a reduced form bargaining model, but it can also be considered somewhat differently. That is, when trust is the issue to be solved, there are several solutions to conflict. Concessions are one way in which countries can signal their commitment.  By modeling the conflict in such a way we can also gain insight as to how a third party can ameliorate trust issues inherent to the conflict. %might want to cite fearon and bargaining lit here
%might want to eventually add the stochastic type switch- or at least talk about how that changes analysis

@@

Here $T>0$ represents the benefit from a country's negotiating partner playing Trust. $D>0$ represents the damages incurred when a country's negotiating partner plays Fight, while $W>0$ is the additional benefit a country receives when it plays Fight itself. We will refer to the stage-game Nash Equilibrium of (Fight, Fight) as ``War.'' We will refer to (Trust,Trust) as ``Peace.'' In line with the assumption that the countries play a Prisoner's Dilemma in the stage game, we assume $T>W-D$. This also ensures that the joint payoffs from ``Peace'' are the highest among all the stage-game action profiles.

@@

We will use grim trigger punishments throughout: that is, (Fight, Fight) will be played forever if Fight is played by any party in any round. Because we use grim trigger punishments, there are only 4 interesting paths of play after concessions are given:
1) Both countries play Trust always; 
2) Both play Fight always; 
3) Country 1 plays Fight in round 1 while country 2 plays Trust, which is followed by both countries playing Fight from round 2 onwards; and 
4) Country 2 plays Fight in round 1 while country 1 plays Trust, which is followed by both countries playing Fight from round 2 onwards.

Equilibrium payoffs beginning from round 1 are now easily described by noting only the collective first round behavior. Let $X_{ij}$ represent the sum of discounted payoffs for country $i$ where subscripts represent first round strategies of that country. If both countries play T in round 1, the corresponding payoff to country $i$ is represented as $ X_{TT}=\frac {T}{1- \delta_i}$. Likewise $X_{TF}=({-D}+\frac{\delta_i(W-D)}{1- \delta_i} )$, $ X_{FF}=\frac{W-D}{1- \delta_i}$, and  $X_{FT}=T+W +\frac{\delta_i(W-D)}{1- \delta_i}$. Note that because types have different discount factors, the payoffs differ by type, and when ambiguous will be super-scripted to indicate the type, e.g. $X_{FT}^h$ represents payoffs for the high type of country 1 when country 1 fights and country 2 trusts in the first period.

@@

\section{Analysis of Benchmark Cases}
\label{sec:bench}
When countries are faced with uncertainty about each other's motives, they can proceed in several different ways, including giving concessions in order to signal their willingness and ability to make peace. We will examine three types of equilibria.\footnote{If the discount factor is high enough, there are many types of equilibria. Consider for instance oscillating every other period between (Trust,Trust) and (Fight,Fight) with a grim trigger ``War'' threat. In this paper, we are not interested in these equilibria because they are not as realistic for applications to conflict scenarios. These are also always payoff dominated by at least one of the other equilibria discussed and thus 	are not attractive.} The Appendix provides a thorough analysis of the various equilibria as well as most proofs.

\subsection{Pooling equilibrium}
In the \emph{pooling equilibrium}, neither negotiating party gives concessions in period 0, then both parties play Fight in every period beginning with period 1. Countries continue to doubt, doing nothing to bridge the credibility gap, and thus not cooperating. This \emph{pooling equilibrium} is likely to happen when countries are in long-set patterns of distrust. Active fighting is not necessary; remaining outside of a state of peace is all that is required.  See the Appendix for a formal characterization of this type of equilibrium.

@@

Without a third party, there is little hope of overcoming mistrust. These long lasting stalemates are the types of conflict which are most appropriate for mediator involvement. In Section~\ref{sec:med}, we will present a mediation mechanism that illustrates how mediation can help achieve peace in this context. In some sense, these are the cases that are most interesting because the international community might play a positive role.

\subsection{No-Concessions Separating Equilibrium}
\label{sec:ncse}
The other two equilibria on which we focus are separating equilibria in which low types and high types have different strategies. If $\delta_i$ is high enough for both countries, the ``Peace'' outcome can be sustained with no concessions and a grim trigger punishment threat. We call this a \emph{no-concessions separating equilibrium}. In this equilibrium, no concessions are given by either type in period 0. In period 1, high types play Trust while low types play Fight. So long as there are two high types as revealed by their period 1 play, both countries play Trust in all following periods. Otherwise, both countries play Fight in all following periods.

@@

\begin{theorem}
\emph{No Concessions Separating Equilibrium}\\
	Assume $\delta_h \geq \frac{W}{(1-p)W + p(T+D)}=\de^{nc}$. A subgame perfect Nash equilibrium exists in which low types give no concessions in period 0 and always play Fight and the high types give no concessions in period 0 and play Trust unless the other country defects to Fight.
	\label{theorem:1}
\end{theorem}
\emph{Proof}: See Appendix.
\\
Here, as in each further iteration of the model, we define a high type as a country who prefers to play the ``Peace'' equilibrium, i.e. $\delta_h \geq \delta^{nc}$. Any country with $\delta_l < \delta^{nc}$ is a low type. This means that two high types---and only two high types---can sustain the ``Peace'' outcome in a simple repeated game.

In this equilibrium, two mutually distrustful countries \emph{behave} as if they trust each other, taking a risk. This generally will happen when the result of trusting and then being betrayed is not catastrophic, or the odds of such a scenario are small. This is representative of a case where countries have suspicions about each other, but proceed anyway.

Examples of a \emph{no-concessions separating equilibrium} are most likely to be those where countries have traditionally had good relations but something has recently come in the way. Certain relationships between the USA and countries that its WikiLeaks alienated -- such as Italy -- would be an example. Italy essentially laughed off the criticisms and reassured the USA of its commitment to their relationship. Warsaw Pact countries generally allowed the USSR similar leeway; it was better to have occasionally imperfect cooperation than conflict. Some events can disrupt trust without completely derailing cooperation.

@@

%Though it will not be modeled here, a stochastic component -- when countries occasionally unilaterally defect -- can explain why powerful countries like the USSR more often ``get away'' with cheating: the payoffs to the relationship are not symmetric. Losing the USA or USSR as an ally is much more devastating for a satellite country than vise versa, therefore big countries can have their alliances and take advantage of them too.
\subsection{Concessions Separating Equilibrium}
\label{sec:cse}
The other separating equilibrium of interest involves distrustful countries using concessions as a way of building trust. We call this the \emph{concessions separating equilibrium}. In this equilibrium, the cost of giving a concession can be worth the investment if the other country also cooperates. It is also not as risky as the \emph{no-concession separating equilibrium}. In this equilibrium, concessions are given before more substantive actions and thus remove the possibility of playing one of the cheating corners of the Prisoner's Dilemma. Concessions are a costly signal that the other country is serious about cooperating and can be sued to avoid continual distrust. That is, high types make use of concessions to identify themselves as trustworthy partners in hopes of establishing peace.

@@

Details of the analysis are in the Appendix. We begin by characterizing the concession (or \emph{gift}) giving behavior of the low type. We then present the patience threshold and gift giving behavior of the high type.

@@

A large probability of a high type makes the \emph{no-concessions separating equilibrium} very attractive because the risk of ending up in the cheating corner of the Prisoner's Dilemma is low and no concessions have to be made. Conversely, for smaller values of $p$, high types are better off paying the gift to avoid the possibility of being taken advantage of by a low type.

@@

\subsection{Concessions with Material Value}
We now model concessions to have real future material value that can either help or harm the giver. To be clear, in Sections~\ref{sec:model} and \ref{sec:bench}, the gift only benefits the party who receives it in the period in which it is received, that is, Period 0. In this section, we assume that there is an additional value from the gift in each period that the repeated game is played, that is, Period 1 and onward.

The intuition for this addition to the model is that the country that receives the concessions can use them to build up their military or to build up their civil society. Using concessions to build military strength hurts the giver of the concessions in war; using the concessions to strengthen civil society helps the giver in peace. The Golan heights is such an example: if Syria is an ally of Israel, the Golan Heights is a reasonable concession to make, but if Syria later turns out to betray a trust, surrender of the strategic land would be disastrous for Israel. %Another example is ADD more examples

@@

Because $\al_i$ enters into the payoffs in a linear fashion, the choice that maximizes welfare in equilibrium is $\alpha=1$ for high types since a concession would only come from another high type in equilibrium, and thus only the ``Peace'' outcome will be played. Conversely, in equilibrium, low types only participate in the ``War'' equilibrium and thus maximize their payoffs by choosing $\alpha=0$.

@@

$\de_h$ is not involved in any of these calculations, so we can see that for any $\de_h \in [.5\overline{3},1)$.\footnote{Recall that we restrict $p<1$ since there is no imperfect information when only one type is possible.} For this range of $\de_h$, separating through concessions is possible when concessions have no material value but is not possible when they do have future material value. That is, future material value destroys the opportunity for peace. \hfill $\blacksquare$

@@

We implement this assumption by allowing countries to choose a scalar $e \in[0,1]$ by which to multiply their given concession $g$. $C(g)= g$ as before, but while the receiving country incurs the full cost of the concession $g$, the future benefit to the recipient is now only $eg$.\footnote{More generally, we could think of $e$ as a function $E(g)$ instead of a simple scalar transformation. Without further restriction, $E(\cdot)$ could increase the value of the concession.} Table 3 provides the modified payoffs in the repeated game.

@@

The intuition for Lemma~\ref{lemma:efficiency} is that future material value of concession helps you when your negotiating partner turns out to be a high type, and this happens $p$ of the time. On other other hand, the material value of your concession \emph{hurts} you when your negotiating partner turns out to be a low type. Similarly, the larger are the benefits from Trust, the larger the welfare gain when the negotiating partner is a high type and the larger are the damages from Fight, the larger is the welfare loss when the negotiating partner turns out to be a low type.

Lemma~\ref{lemma:efficiency} implies that the high types will either prefer $e=1$ or $e=0$. As we turn to our result about incentive compatibility when concessions have material value, we focus on these two cases as the most focal and easiest upon which to coordinate.\footnote{For some parameter values, we cannot rule out that a \emph{concessions separating equilibrium} with $0<e<1$ exists and improves upon high type welfare when high types are patient enough to separate when $e=0$ but not when $e=1$. This is not at all obvious though, especially because increasing $e$ from $0$ can tighten the high type's incentive compatibility constraint.}

@@

This means that there are parameters under which ``Peace'' cannot be achieved in a \emph{concessions separating equilibrium with efficient gifts} but ``Peace'' can be achieved in a \emph{concessions separating equilibrium with inefficient gifts}. This is part of the substance of Theorem~\ref{theorem:4}.

\begin{theorem}
		Assume the high types are not patient enough to separate without concessions and that concessions have material value that can be destroyed. There are parameters under which the optimal equilibrium is a separating equilibrium in which concessions are inefficient.
	\label{theorem:4}
\end{theorem}
\emph{Proof}: We invoke Lemmas~\ref{lemma:efficiency} and \ref{lemma:e-patient} as two different reasons for which the parties would choose the \emph{concessions separating equilibrium with inefficient gifts}. First, they may be patient enough to separate when gifts are inefficient but not when gifts are efficient (Lemma~\ref{lemma:e-patient}). In this case, the \emph{concessions separating equilibrium with efficient gifts} is not available to them.

@@

Inefficient concessions are optimal when there's a good chance of facing a low type who will use the concession against you (i.e. $p$ is small) as well as when the benefits from ``Peace'' are low relative to the damages from ``War.'' This finding of inefficiency is for profoundly different reasons than the literature related to Camerer (1988) or Prendergast and Stole (2001).\footnote{Since our initial setup is essentially identical to Camerer's, inefficiency of  pre-play communication can also be found in our model. With a slight behavioral modification, the reasoning of Prendergast and Stole (2001) can also be had: if a country suffers regret after efficiently giving to a type different than their own, there is a premium for guessing correctly. In Prendergast and Stole the two types are equal except in labeling; in our model, high types and low types are used so there is an incentive for low types to try to pose as high types. However, putting this difference aside, the inefficiency result can be viewed as similar to that in Prendergast and Stole (2001). That is, if countries suffer a ``mismatching'' cost due to giving to the wrong type, they will be hesitant to do so and may offer inefficient concessions.} Instead of being in reaction to a behavioral regret for mismatching or serving as a way to discourage low types from entering the game altogether, here making concessions inefficient is a device to prevent low types from using those concessions against the giver of the concessions in the future.

@@

In Sections~\ref{sec:bench} and \ref{sec:IC}, the same concession must be sent to both types since types are private information. Importantly, here the mechanism $M$ can differentiate based on recipient's type. Otherwise, the setup is the same as in Section~\ref{sec:matval} where concessions have future material value that can optionally be destroyed. As usual, truthful self-revelation requires that the incentive compatibility constraint for each type must be satisfied.

@@

\begin{lemma}
	A manipulative mediator cannot facilitate the most efficient concessions, i.e. $e=1$ is not incentive compatible under mechanism $M$. 
	\label{lemma:med}
\end{lemma}
\emph{Proof}: See Appendix.

@@

\begin{theorem}
	Assume that concessions have future material value.
	\begin{enumerate}
		\item[(a)] If future material value can be destroyed, a mediator can be used to eliminate inefficient concessions and improve welfare.
		\item[(b)] If future material value cannot be destroyed, a mediator can be used to expand the parameter space over which peace can be achieved.
	\end{enumerate}
	\label{theorem:5}
\end{theorem}
\emph{Proof:} (a) Under $M$, the low type incentive compatibility constraint can bind without having high types give a concession to low types. There is no need then to give inefficient concessions: high types will invest in civil society and improve the giver's welfare and low types will never receive the concession. %The equilibrium separating concession $g^{*M}$ will be larger than $g^*$ in this case because the low type who reports truthfully receives a concession with zero probability: 
%\be
%g^{*M}= \frac{p(X_{FT}^l-X_{FF}^l)}{1-p}
%\ee
High type utility under $M$ is 
\be
U_h^M= pX_{TT}+(1-p)X_{FF} = p \left( \frac{T(1+g^{*M})}{1-\de_h} \right) + (1-p) \left(\frac{W-D}{1-\de_h} \right)
\ee
because high types exchange concessions of the same value. Welfare, defined as high-type utility, is improved because no transfer is made to the low types and no future damage is caused in the case of such a transfer.\footnote{Low types IR constraint still binds and $U_l=X_{FF}$.}

@@

The low type leader can bluff in the bilateral setting but not when a mediator is present. As per Theorem~\ref{theorem:4}, inefficiency is due to the high types giving an inefficient gift in order to separate while still being able to prevent low types from using concessions against the high types. But because the benefits of giving a concession differ between the types (low types get one period of cheating payoff, while high types get future peace), the mediator is able to mandate concessions only between declared high types. This means there is no longer a need for inefficient concessions since concessions will never be used against the giver. Thus, a mediator is able to remove inefficiency in concession giving and bring willing parties to the table to give the gifts necessary to achieve peace.

@@

 \noindent{\emph{Proof of Theorem \ref{theorem:1}}:\\
Playing Trust in the first period and thereafter as long as Trust has always been played will be incentive compatible for high types if the expected payoffs from this strategy are greater than the payoffs from playing Fight in the first period and then every period thereafter by the one-shot deviation principle. If both countries play Trust in a period, a ``Peace'' equilibrium takes place and both countries continue to play Trust in every period with the grim trigger threat of the ``War'' equilibrium. Low types play Fight in the first period, so a high type that observed Fight being played would respond by playing Fight in period 2 and in all future periods. A country plays Trust if and only if the following condition is satisfied:
\begin{equation*}
  p\left(\frac{T}{1- \delta_i}\right) +(1-p)\left({-D}+\frac{\delta_h(W-D)}{1- \delta_i} \right)\geq p \left(T + W + \frac{\de_h}{1-\de_h}(W-D)\right) +(1-p)\frac{W-D}{1- \delta_h}
\end{equation*}
Simplifying, we have
%\be
%  p\left(\frac{T}{1- \delta_h}\right) \geq p \left(T + W + \frac{\de_h}{1-\de_h}(W-D)\right) +(1-p)W
%\ee
\begin{equation*}
  pT \geq p \left((1-\de_h)(T + W) + \de_h (W-D)\right) + (1-\de_h)(1-p)W
\end{equation*}
\begin{equation*}
  pT \geq pT -p\de_hT +pW -p\de_h W + p \de_h W -p \de_h D  + W -\de_h W -pW + p\de_hW
\end{equation*}
\begin{equation*}
  p\de_hT +p \de_h D + \de_h W - p\de_hW  \geq W 
\end{equation*}
\be
	\delta_h \geq \frac{W}{(1-p)W + p(T+D)}
	\label{eq:ncse}
\ee
